---
title: "Define specific insertions"
output: html_document
date: "2023-07-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read blast results

```{r read_blast}
#read blast output
mm1s_match <- read.table("result/assembly_blast/MM1S_matchet_assembly_777_blast.txt",colClasses = c(rep("character", times = 2), rep("numeric", times = 10)))
mia_match <- read.table("result/assembly_blast/MIA_matched_P201812-2_blast.txt",colClasses = c(rep("character", times = 2), rep("numeric", times = 10)))
mm1s_mismatch <- read.table("result/assembly_blast/MM1S_mismatch_assembly_777_blast.txt",colClasses = c(rep("character", times = 2), rep("numeric", times = 10)))
mia_mismatch <- read.table("result/assembly_blast/MIA_mismatch_P201812-2_blast.txt",colClasses = c(rep("character", times = 2), rep("numeric", times = 10)))
mm1s_cell_line <- read.table("result/assembly_blast/MM1S_cell_line_assembly_777_blast.txt",colClasses = c(rep("character", times = 2), rep("numeric", times = 10)))
mia_cell_line <- read.table("result/assembly_blast/MIA_cell_line_P201812-2_blast.txt",colClasses = c(rep("character", times = 2), rep("numeric", times = 10)))

#define column name
colnames(mm1s_match) <- c("query", "subject", "%_identity", "alignment_length", "mismatches", "gap_opens",
                          "q_start", "q_end", "s_start", "s_end", "evalue", "bit_score")
colnames(mia_match) <- c("query", "subject", "%_identity", "alignment_length", "mismatches", "gap_opens",
                          "q_start", "q_end", "s_start", "s_end", "evalue", "bit_score")
colnames(mm1s_cell_line) <- c("query", "subject", "%_identity", "alignment_length", "mismatches", "gap_opens",
                         "q_start", "q_end", "s_start", "s_end", "evalue", "bit_score")
colnames(mia_cell_line) <- c("query", "subject", "%_identity", "alignment_length", "mismatches", "gap_opens",
                         "q_start", "q_end", "s_start", "s_end", "evalue", "bit_score")
colnames(mm1s_mismatch) <- c("query", "subject", "%_identity", "alignment_length", "mismatches", "gap_opens",
                             "q_start", "q_end", "s_start", "s_end", "evalue", "bit_score")
colnames(mia_mismatch) <- c("query", "subject", "%_identity", "alignment_length", "mismatches", "gap_opens",
                             "q_start", "q_end", "s_start", "s_end", "evalue", "bit_score")
```

Statistics for insertion

```{r blast}

# define number of insertions
####
paste("# inst mm1s_match = ", length(mm1s_match$query))
paste("# unique inst mm1s_match = ", length(unique(mm1s_match$query)))
paste("# inst mm1s_mismatch = ", length(mm1s_mismatch$query))
paste("# unique inst mm1s_mismatch = ", length(unique(mm1s_mismatch$query)))
paste("# inst mm1s_cell_line = ", length(mm1s_cell_line$query))
paste("# unique inst mm1s_cell_line = ", length(unique(mm1s_cell_line$query)))
paste("# inst mia_match = ", length(mia_match$query))
paste("# unique inst mia_match = ", length(unique(mia_match$query)))
paste("# inst mia_mismatch = ", length(mia_mismatch$query))
paste("# unique inst mia_mismatch = ", length(unique(mia_mismatch$query)))
paste("# inst mia_cell_line = ", length(mia_cell_line$query))
paste("# unique inst mia_cell_line = ", length(unique(mia_cell_line$query)))

# visualization
combined_aln_length <- data.frame(alignment_length = NA,
                                  group = NA)
combined_aln_length <- rbind(combined_aln_length, data.frame(alignment_length =mm1s_match$alignment_length,
                                                             group = "mm1s_match"))
combined_aln_length <- rbind(combined_aln_length, data.frame(alignment_length =mm1s_cell_line$alignment_length,
                                                             group = "mm1s_cell_line"))
combined_aln_length <- rbind(combined_aln_length, data.frame(alignment_length =mia_match$alignment_length,
                                                             group = "mia_match"))
combined_aln_length <- rbind(combined_aln_length, data.frame(alignment_length =mia_cell_line$alignment_length,
                                                             group = "mia_cell_line"))
combined_aln_length <- rbind(combined_aln_length, data.frame(alignment_length =mm1s_mismatch$alignment_length,
                                                             group = "mm1s_mismatch"))
combined_aln_length <- rbind(combined_aln_length, data.frame(alignment_length =mia_mismatch$alignment_length,
                                                             group = "mia_mismatch"))
combined_aln_length <- na.omit(combined_aln_length)

boxplot(combined_aln_length$alignment_length ~ combined_aln_length$group, main = "Alignment length",
        xlab = "group", ylab = "Alignment length")
```

Define unique hits for each sample, compared with cell lines.
Unique hits of mm1s_match compared with mm1s_cell_line \ 
Unique hits of mia_match compared with mia_cell_line \ 
Unique hits of mm1s_mismatch_uniq compared with mm1s_cell_line \ 
Unique hits of mia_mismatch_uniq compared with  mia_cell_line \ 
 

```{r blast_uniq_hits}
## define unique hits for mm1s_match compared with mm1s cell line

mm1s_match_uniq <- mm1s_match[-which(mm1s_match$subject %in% intersect(mm1s_match$subject, mm1s_cell_line$subject)),]
hist(mm1s_match_uniq$alignment_length, main = "mm1s_match unique hits", xlab = "Alignment length")

## define unique hits for mm1s_match compared with mia cell line
mia_match_uniq <- mia_match[-which(mia_match$subject %in% intersect(mia_match$subject, mia_cell_line$subject)),]
hist(mia_match_uniq$alignment_length, main = "mia_match unique hits", xlab = "Alignment length")

## define unique hits for mm1s_mismatch compared with  mm1s cell line

mm1s_mismatch_uniq <- mm1s_mismatch[-which(mm1s_mismatch$subject %in% intersect(mm1s_mismatch$subject, mm1s_cell_line$subject)),]
hist(mm1s_mismatch_uniq$alignment_length, main = "mm1s_mismatch unique hits", xlab = "Alignment length")

## define unique hits for mia_mismatch compared with  mia cell line
mia_mismatch_uniq <- mia_mismatch[-which(mia_mismatch$subject %in% intersect(mia_mismatch$subject, mia_cell_line$subject)),]
hist(mia_mismatch_uniq$alignment_length, main = "mia_mismatch unique hits", xlab = "Alignment length")
```
Define statistics for unique samples

```{r unique_sample_stat}
##### define number of hits
#match
# paste("# hits mm1s_match = ", length(mm1s_match_uniq$subject))
paste("# unique hits mm1s_match = ", length(unique(mm1s_match_uniq$subject)))
mm1s_match_aln_length_stat <- as.matrix(summary(mm1s_match_uniq$alignment_length))
colnames(mm1s_match_aln_length_stat) <- "mm1s_match_uniq"
mm1s_match_aln_length_stat


paste("# hits mia_match = ", length(mia_match_uniq$subject))
paste("# unique hits mia_match = ", length(unique(mia_match_uniq$subject)))
mia_match_aln_length_stat <- as.matrix(summary(mia_match_uniq$alignment_length))
colnames(mia_match_aln_length_stat) <- "mia_match_uniq"
mia_match_aln_length_stat

#mismatch
paste("# hits mm1s_mismatch = ", length(mm1s_mismatch_uniq$subject))
paste("# unique hits mm1s_mismatch = ", length(unique(mm1s_mismatch_uniq$subject)))
mm1s_mismatch_aln_length_stat <- as.matrix(summary(mm1s_mismatch_uniq$alignment_length))
colnames(mm1s_mismatch_aln_length_stat) <- "mm1s_mismatch_uniq"
mm1s_mismatch_aln_length_stat

paste("# hits mia_mismatch = ", length(mia_mismatch_uniq$subject))
paste("# unique hits mia_mismatch = ", length(unique(mia_mismatch_uniq$subject)))
mia_mismatch_aln_length_stat <- as.matrix(summary(mia_mismatch_uniq$alignment_length))
colnames(mia_mismatch_aln_length_stat) <- "mia_mismatch_uniq"
mia_mismatch_aln_length_stat

rm(mm1s_match_aln_length_stat, mia_match_aln_length_stat, mm1s_mismatch_aln_length_stat, mia_mismatch_aln_length_stat)

```


Define unique hits for match samples, compared with mistatched and cell lines

```{r blast_uniq_hits_1}
## compare unique mm1S match with mm1s mismatch, define mm1S match unique
mm1s_match_mm1s_mismatch_uniq <- mm1s_match_uniq[-which(mm1s_match_uniq$subject %in% intersect(mm1s_match_uniq$subject, mm1s_mismatch$subject)),]

## compare unique mia match with mia mismatch, define mia match unique
mia_match_mia_mismatch_uniq <- mia_match_uniq[-which(mia_match_uniq$subject %in% intersect(mia_match_uniq$subject, mia_mismatch$subject)),]

print("MM1S")
# number of hits mm1s
paste("mm1s_match_mm1s_mismatch_uniq # hits = ", length(mm1s_match_mm1s_mismatch_uniq$subject))
paste("mm1s_match_mm1s_mismatch_uniq # unique hits = ",length(unique(mm1s_match_mm1s_mismatch_uniq$subject)))
# number of hits > 80% intensity mm1s
paste("mm1s_match_mm1s_mismatch_uniq # hits (>80) = ", length(mm1s_match_mm1s_mismatch_uniq$subject[mm1s_match_mm1s_mismatch_uniq$`%_identity` > 80 ]))
paste("mm1s_match_mm1s_mismatch_uniq # unique hits (>80) = ", length(unique(mm1s_match_mm1s_mismatch_uniq$subject[mm1s_match_mm1s_mismatch_uniq$`%_identity` > 80 ])))
# number of hits > 90% intensity mm1s
paste("mm1s_match_mm1s_mismatch_uniq # hits (>90) = ", length(mm1s_match_mm1s_mismatch_uniq$subject[mm1s_match_mm1s_mismatch_uniq$`%_identity` > 90 ]))
paste("mm1s_match_mm1s_mismatch_uniq # unique hits (>90) = ", length(unique(mm1s_match_mm1s_mismatch_uniq$subject[mm1s_match_mm1s_mismatch_uniq$`%_identity` > 90 ])))
# number of hits == 100% intensity mm1s
paste("mm1s_match_mm1s_mismatch_uniq # hits (100%) = ", length(mm1s_match_mm1s_mismatch_uniq$subject[mm1s_match_mm1s_mismatch_uniq$`%_identity`  == 100 ]))
paste("mm1s_match_mm1s_mismatch_uniq # unique hits (100%) = ", length(unique(mm1s_match_mm1s_mismatch_uniq$subject[mm1s_match_mm1s_mismatch_uniq$`%_identity` == 100])))

print("MIA")
# number of hits mia
paste("mia_match_mia_mismatch_uniq # hits = ", length(mia_match_mia_mismatch_uniq$subject))
paste("mia_match_mia_mismatch_uniq # unique hits = ",length(unique(mia_match_mia_mismatch_uniq$subject)))
# number of hits > 80% intensity mia
paste("mia_match_mia_mismatch_uniq # hits (>80) = ", length(mia_match_mia_mismatch_uniq$subject[mia_match_mia_mismatch_uniq$`%_identity` > 80 ]))
paste("mia_match_mia_mismatch_uniq # unique hits (>80) = ", length(unique(mia_match_mia_mismatch_uniq$subject[mia_match_mia_mismatch_uniq$`%_identity` > 80 ])))
# number of hits > 90% intensity mia
paste("mia_match_mia_mismatch_uniq # hits (>90) = ", length(mia_match_mia_mismatch_uniq$subject[mia_match_mia_mismatch_uniq$`%_identity` > 90 ]))
paste("mia_match_mia_mismatch_uniq # unique hits (>90) = ", length(unique(mia_match_mia_mismatch_uniq$subject[mia_match_mia_mismatch_uniq$`%_identity` > 90 ])))
# number of hits == 100% intensity mia
paste("mia_match_mia_mismatch_uniq # hits (100%) = ", length(mia_match_mia_mismatch_uniq$subject[mia_match_mia_mismatch_uniq$`%_identity` == 100 ]))
paste("mia_match_mia_mismatch_uniq # unique hits (100%) = ", length(unique(mia_match_mia_mismatch_uniq$subject[mia_match_mia_mismatch_uniq$`%_identity` == 100])))

```

Save unique hits and insertions

```{r save_uniq}
# MM1S
write.table(paste(">", unique(mm1s_match_mm1s_mismatch_uniq$subject), sep = ""), "result/assembly_blast/MM1S_match_unique_contigs.txt", quote = F, row.names = F,col.names = F, sep = "\t")
write.table(mm1s_match_mm1s_mismatch_uniq, "result/assembly_blast/MM1S_match_unique_contigs_blast_result.txt")

#MIA
write.table(paste(">", unique(mia_match_mia_mismatch_uniq$subject), sep = ""), "result/assembly_blast/MIA_match_unique_contigs.txt",quote = F, row.names = F,col.names = F, sep = "\t")
write.table(mia_match_mia_mismatch_uniq, "result/assembly_blast/MIA_match_unique_contigs_blast_result.txt")
```
